Modern Statistical Methods --- Example Sheet 1

### 2020-10-27 distribution_theory/C2.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

	Fix $\phi, \psi \in \DD(\RR^n)$, we see from our earlier shown properties that 
	\begin{align*}
	(u * v) * (\phi * \psi) &= u * \qty(v * (\phi * \psi)) = u * ((v * \phi) * \psi) = u * (\psi * (v * \phi)) = (u * \psi) * (v * \phi).
	\end{align*}
	If we interchange $u$ and $v$ in the above, that is equivalent to interchanging $\phi$ and $\psi$, which we know must yield the same result. This shows $u * v$ and $v * u$ agree on $\phi * \psi$ for all $\phi, \psi \in \DD(\RR^n)$, and from this it follows that $u * v = v * u$ (\TODO why?).
\end{proof}

>>>>

	Fix $\phi, \psi \in \DD(\RR^n)$, we see from our earlier shown properties that 
	\begin{align*}
	(u * v) * (\phi * \psi) &= u * \qty(v * (\phi * \psi)) = u * ((v * \phi) * \psi) = u * (\psi * (v * \phi)) = (u * \psi) * (v * \phi).
	\end{align*}
	If we interchange $u$ and $v$ in the above, that is equivalent to interchanging $\phi$ and $\psi$, which we know must yield the same result. This shows $u * v$ and $v * u$ agree on $\phi * \psi$ for all $\phi, \psi \in \DD(\RR^n)$. Defining $E = u * v - v * u$, we find that $0 = E * (\phi * \psi) = (E * \phi) * \psi$ for all $\phi, \psi \in \DD(\RR^n)$, so $E * \phi = 0$ for all $\phi \in \DD(\RR^n)$, so $E = 0$. 
\end{proof}

### 2020-10-27 distribution_theory/distribution_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}
\newcommand{\DD}{\Cal D}

@@

\begin{solution}
	Let $\phi \in \DD(X)$. 
	\begin{enumerate}[(a)]
		\item If $u' = 1$ then we find 
		\[
		\int_\RR \phi(x) \dd{x} = \ang{1, \phi} = \ang{u', \phi} = -\ang{u, \phi'},
		\]
		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$. 
	\end{enumerate}
\end{solution}
\end{document}

>>>>

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Cal E}
\newcommand{\DD}{\Cal D}
\newcommand\ZZ{\Bb Z}

@@

\begin{solution}
	Let $\phi \in \DD(X)$. 
	\begin{enumerate}[(a)]
		\item If $u' = 1$ then we find 
		\[
		\int_\RR \phi(x) \dd{x} = \ang{1, \phi} = \ang{u', \phi} = -\ang{u, \phi'},
		\]
		so for any $c \in \RR$ we find by partial integration $\ang{u, \phi'} = - \int_\RR\phi(x) \dd{x} = \int_\RR (x + c) \phi'(x) \dd{x}$. From this we deduce that $u = x + c$ for some $c$.

		\item If $xu' = \delta_0$ then we have
		\[
		\phi(0) = \ang{\delta_0, \phi} = \ang{xu', \phi} = \ang{u', x\phi} = \ang{u, -(x\phi)'} = \ang{u, -x\phi' - \phi} %=  \ang{u, -x\phi'} + \ang{u, -\phi}. 
		\]
		Note that the above is satisfied for $u = - \delta_0 + c$ for any constant $c$. \TODO is this the most general solution?

		\item Since $e^{2\pi i x} = 1 \iff x \in \ZZ$, intuitively it must be the case that $u'$ is 0, except ``on $\ZZ$'', whatever that may mean. Therefore, we guess that, for any sequence $(\alpha_n)_{n \in \ZZ} \subseteq \CC$ and constant $c \in \CC$, the map \[u = c + \sum_{n \in \ZZ} \alpha_n \ind_{x \geq n}.\] We compute the derivative of $u$. It is easily seen that $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$ (the infiniteness of the sum does not pose a problem since the test functions are compactly supported, so $\ang{u, \phi}$ will always be a finite sum). 
%		We have for arbirary $\phi$ that
%		\begin{align*}
%		\ang{u', \phi} &= - \ang{u, \phi'} = -\int_\RR \sum_{n \in \ZZ} \qty(\alpha_n \ind_{x \geq n}) \phi'(x) \dd{x}	\overset\star= - \sum_{n \in \ZZ} \alpha_n \int_n^\infty \phi'(x) \dd{x} = \sum_{n \in \ZZ} \alpha_n \phi(n),	
%		\end{align*}
%	where $\star$ follows from the fact that the sum is finite since $\phi'$ is compactly supported. 
%	Therefore, we can conclude $u' = \sum_{n \in \ZZ} \alpha_n \delta_n$, and indeed we find
From this, we see that
	\[
	\ang{(e^{2\pi i x} - 1) u', \phi} = \ang{u', (e^{2\pi i x} -1 )\phi} = \sum_{n \in \ZZ} \alpha_n (e^{2\pi i n} - 1) \phi(n) = 0,
	\]
	so $u$ satisfies the equation. \TODO Why is this the most general solution? Intuitively clear, but how to make this rigorous?
	\end{enumerate}
\end{solution}

\begin{question}
	Define the distribution $u \in \DD'(\RR^2)$ by the locally integrable function $u(x, y) = \ind_{x \geq y}$. Show that $\pt_x^2   u - \pt_y^2 u = 0$ in $\DD'(\RR^2)$. Can you give a physical interpretation of this result?
\end{question}

\begin{proof}
	Let $f \in \DD(\RR^2)$, then we have
	\begin{align*}
		\ang{\pt_x^2 u - \pt_y^2 u, f} &= \ang{\pt_x^2 u, f}- \ang{\pt_y^2 u, f} = \ang{u, \pt_x^2 f} - \ang{u, \pt_y^2 f} = \ang{u, \pt_x^2 f - \pt_y^2 f} \\
		&\overset\star= \int_{-\infty}^\infty \int_{y}^\infty \pt_x^2 f(x, y) \dd{x}\dd{y} - \int_{-\infty}^\infty \int_{-\infty}^x \pt_y^2 f(x, y) \dd{y} \dd{x} \\
		&= - \int_{-\infty}^\infty \pt_x f(y, y) \dd{y} - \int_{-\infty}^\infty \pt_y f(x, x) \dd{x}
		\\
		&= - \int_{-\infty}^\infty (\pt_xf + \pt_yf)(x, x) \dd{x}.
	\end{align*}
Here, $\star$ follows from Fubini's theorem. 
Define $g(x) = f(x, x)$, then it is easily seen that $g'(x) = \pt_x f(x, x) + \pt_y f(x, x)$, so we find that
\[
\ang{\pt_x^2 u - \pt_y^2 u, f} = 
- \int_{-\infty}^\infty g'(x) \dd{x} = \lim_{x\to -\infty} g(x) - \lim_{x\to\infty} g(x) = 0 - 0 = 0. 
\]
\end{proof}

This shows that $\pt_x u - \pt_y u = 0$, or equivalently, that $u$ satisfies the wave equation. 
\end{document}

### 2020-10-27 inverse_problems/inverse_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

@@

\newcommand\XX{\Cal X}
\newcommand\YY{\Cal Y}
\newcommand\UU{\Cal U}
\newcommand\VV{\Cal V}
\newcommand\LL{\Cal L}
\newcommand\KK{\Cal K}
\newcommand\BB{\Cal B}
% \newcommand\LLL{\Bb L}
\DeclareMathOperator\Dom{\Cal D}
\DeclareMathOperator\Nul{\Cal N}
\DeclareMathOperator\Ran{\Cal R}

@@

\begin{question}
	For $\Omega = [0, 1]^2$ and $\XX \in L^2(\Omega)$, we consider the integral operator $A \colon \XX \to \XX$ with
	\[
	(Au)(y) \ceq \int_\Omega k(x, y) u(x) \dd{x},
	\]
	for $k \in L^2(\Omega \times \Omega)$.  Show that
	\begin{enumerate}[(a)]
		\item $A$ is linear with respect to $u$,
		\item $A$ is a bounded linear operator, i.e.\ $\norm{Au}_\XX \leq \norm{A}_{\LL(\XX, \XX)} \norm{u}_\XX$. Give also an estimate for $\norm{A}_{\LL(\XX, \XX)}$,
		\item the adjoint $A^*$ is given via
		\[
		(A^* v)(y) = \int_\Omega k(y, x) v(x) \dd{x}.
		\]
		\item $A$ is a compact operator, i.e.\ $A \in \KK(\XX, \XX)$.

@@

\begin{solution}
	Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products.

@@

\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$. Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2(\XX)$. 
%\[
%(A_n u)(y) \ceq \int_\Omega p_n()
%\]
 \TODO Finish
	\end{enumerate}
\end{solution}

@@

\begin{solution}
	\begin{enumerate}[(a)]
		\item We want to show that we can approximate $f$ by a sequence $(Au_n)$ for some $(u_n) \subseteq L^2[0,1]$. To this end, define for $n \geq 2$
		\[
		u_n(x) =
		\begin{cases}
			0 &\abs{x - \frac12} > \frac1n, \\
			\frac n2 &\abs{x - \frac12} \leq \frac1n. 
		\end{cases}
		\]
		Clearly $u \in L^2[0, 1]$, and we have
		\[
		f_n(y) \ceq (Au_n)(y) = \int_0^y u_n(x) \dd{x} = \begin{cases}
			0 &\text{if $y < \frac12 - \frac1n$,}\\
			\frac n2(y - \frac12 + \frac1n) &\text{if $\frac12 - \frac1n \leq y \leq \frac12 + \frac1n$} \\
			1 &\text{ if $y > \frac12 + \frac1n$}. 
		\end{cases}
		\]

@@

	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
	%Note that 
	%\[
	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
	%\]
	\item To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.

	Note that 
	\begin{align*}
		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
		&= \ang{u, A^* v}
	\end{align*}
where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
\begin{align*}
	(AA^* u)(x) = \int_0^y \int_x^1 u(x) \dd{x} \dd{y}
\end{align*}
	\end{enumerate}
\end{solution}
\end{document}

>>>>

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}
\newcommand\ZZ{\Bb Z}
\newcommand\FF{\Cal F}

@@

\newcommand\XX{\Cal X}
\newcommand\YY{\Cal Y}
\newcommand\UU{\Cal U}
\newcommand\VV{\Cal V}
\newcommand\LL{\Cal L}
\newcommand\KK{\Cal K}
\newcommand\BB{\Cal B}
% \newcommand\LLL{\Bb L}
\DeclareMathOperator\Dom{\Cal D}
\DeclareMathOperator\Nul{\Cal N}
\DeclareMathOperator\Ran{\Cal R}
\DeclareMathOperator\diag{diag}

@@

\begin{mdframed}
		Note: when writing a norm of a vector $v \in V$, I will simply write $\norm{v}$ and not $\norm{v}_V$, unless it is unclear in which space $v$ lives. The same holds for inner products. 	
\end{mdframed}
\begin{question}
	For $\Omega = [0, 1]^2$ and $\XX \in L^2(\Omega)$, we consider the integral operator $A \colon \XX \to \XX$ with
	\[
	(Au)(y) \ceq \int_\Omega k(x, y) u(x) \dd{x},
	\]
	for $k \in L^2(\Omega \times \Omega)$.  Show that
	\begin{enumerate}[(a)]
		\item $A$ is linear with respect to $u$,
		\item $A$ is a bounded linear operator, i.e.\ $\norm{Au}_\XX \leq \norm{A}_{\LL(\XX, \XX)} \norm{u}_\XX$. Give also an estimate for $\norm{A}_{\LL(\XX, \XX)}$,
		\item the adjoint $A^*$ is given via
		\[
		(A^* v)(y) = \int_\Omega k(y, x) v(x) \dd{x}.
		\]
		\item $A$ is a compact operator, i.e.\ $A \in \KK(\XX, \XX)$.

@@

\begin{solution}

@@

\item It is known that for any compact set $X \subseteq \RR^n$, polynomials lie dense in $L^2(X)$. Therefore, there exists a sequence of polynomials $p_n$ such that $p_n \to k$ in $L^2([0, 1]^4)$. It is easily seen that for any polynomial $p$, the operator
\[
(A_p u)(y) \ceq \int_\Omega p(x, y) u(x) \dd{x} 
\] 
has finite rank: let $p(z) = \sum_{\abs{\alpha} \leq n} c_\alpha z^\alpha$ (where $z \in [0, 1]^4$ and $\alpha$ is a multi-index), then we find
\[
(A_p u)(y) = \sum_{\abs{\alpha} \leq n} c_\alpha \int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} y_1^{\alpha_3} y_2^{\alpha_4} u(x) \dd{x} = \sum_{\abs{\alpha} \leq n} c_\alpha \qty(\int_\Omega x_1^{\alpha_1} x_2^{\alpha_2} \dd{x}) y_1^{\alpha_3} y_2^{\alpha_4},
\]
so $A_p u$ lies in the $\Span\qty{y_1^{\alpha_1} y_2^{\alpha_2} \mid \alpha_1 + \alpha_2 \leq n}$, and therefore has finite rank. 
By (b), we find that $\norm{A - A_n} \leq \norm{k - p_n} \to 0$, which shows that $A_n \to A$ in operator norm. We conclude that $A$ is compact. 
%\[
%(A_n u)(y) \ceq \int_\Omega p_n()
%\]
	\end{enumerate}
\end{solution}

@@

\begin{solution}
	\begin{enumerate}[(a)]
		\item We want to show that we can approximate $f$ by a sequence $(Au_n)$ for some $(u_n) \subseteq L^2[0,1]$. To this end, define for $n \geq 2$
		\[
		u_n(x) =
		\begin{cases}
			0 &\abs{x - \frac12} > \frac1n, \\
			\frac n2 &\abs{x - \frac12} \leq \frac1n. 
		\end{cases}
		\]
		Clearly $u \in L^2[0, 1]$, and we have
		\[
		f_n(y) \ceq (Au_n)(y) = \int_0^y u_n(x) \dd{x} = \begin{cases}
			0 &\text{if $y < \frac12 - \frac1n$,}\\
			\frac n2(y - \frac12 + \frac1n) &\text{if $\frac12 - \frac1n \leq y \leq \frac12 + \frac1n$} \\
			1 &\text{if $y > \frac12 + \frac1n$}. 
		\end{cases}
		\]

@@

	% \item From Fourier theory it is known that $\qty{s_n(x) = \sqrt 2\sin(nx) : n \in \NN}$ and $\qty{1} \cup \qty{c_n(x) = \sqrt 2\cos(nx) \mid n \in \NN}$ are orthonormal bases of $L^2[0, 1]$. 
	%Note that 
	%\[
	%(A s_n)(y) = \int_0^y s_n(x) \dd{x} = \sqrt 2 \int_0^y \sin(nx) \dd{x} = \frac{\sqrt 2}{n} \qty(1 - \cos(ny)) = \frac{\sqrt 2}{n} c_0(y) - \frac1n c_n(y) 
	%\]
	\item 	Note that $A$ is compact, since we can write $(Au)(y) = \int_{[0, 1]} k(x, y) u(x) \dd{x}$ with $k(x, y) = \ind_{x \leq y}$.

	To apply the Picard criterion we must find the singular values and right singular vectors of $A$, which are equal to the square roots of the eigenvalues of $AA^*$ and the eigenvectors of $AA^*$.

	\begin{align*}
		\ang{Au, v} &= \int_0^1 (Au)(y)\cdot v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) \dd{x} v(y) \dd{y} \\
		&= \int_0^1 \int_0^y u(x) v(y) \dd{x} \dd{y} \\
		&= \int_0^1 \int_x^1 u(x) v(y) \dd{y} \dd{x} \\ 
		&= \int_0^1 u(x) \int_x^1 v(y) \dd{y} \dd{x} \\
		&= \ang{u, A^* v}
	\end{align*}
where $v(x) = \int_x^1 v(y) \dd{y}$.  Therefore, we find
\begin{align*}
	(AA^* u)(x) = \int_0^x \int_y^1 u(z) \dd{z} \dd{y}.
\end{align*}
Now we solve the eigenequation. Firstly, note that $\Nul(A) = \qty{0}$, so let $\lambda^2 > 0$. Then we have
\begin{align}
	\int_0^x \int_y^1 u(z) \dd{z} \dd{y} &= \lambda^2 u(x) \\
	\int_x^1 u(z) \dd{z} &= \lambda^2 u'(x) \\
	- u(x) &= \lambda^2 u''(x). 
\end{align}
Furthermore, from (1) we infer $u(0) = 0$ while from (2) we infer $u'(1) = 0$. 
The general solution to (3) is given by $u(x) = a \cos(\lambda^{-1} x) + b \sin(\lambda^{-1} x)$. 
Plugging $u(0) = 0$ gives $a = 0$, and $u'(1) = 0$ gives
\[
b\lambda^{-1} \cos(\lambda^{-1}) = 0 \implies b = 0 \text{ or } \lambda^{-1} = \pi\qty(n - \frac12) \text{ for some $n \in \ZZ_{>0}$}.
\]

Our singular values are therefore $\frac{1}{(n - \frac12)\pi}$ ($n \in \ZZ_{>0}$) with right singular vectors $y_n(x) = \sqrt 2\sin((n - \frac12) \pi x)$. 
We now compute
\[
\ang{f, y_n} = \sqrt 2\int_{1/2}^1 \sin(\qty(n - \frac12) \pi x) \dd{x} = ...
\]

\item The Moore-Penrose inverse of $A$ is discontinuous. This can be seen by theorem 2.1.11: in (a) we have constructed a sequence $(f_n) \subseteq \Ran(A)$ that converges, and in (b) we have shown that its limit lies outside of $\Ran(A)$. Therefore, $\Ran(A)$ is not closed, so $A\D$ is discontinuous.

\end{enumerate}
\end{solution}

\begin{question}
	\begin{enumerate}[(a)]
		\item

	Let $m, n \in \NN$ with $m \geq n \geq 2$. Compute the Moore-Penrose inverses of the following matrices:
	\begin{enumerate}[(i)]
		\item $A = (1, 1, \dotsc, 1) \in \RR^{1 \times n}$;
		\item $A = \diag(a_1, \dotsc, a_n) \in \RR^{n \times n}$ with $a_j \in \RR$ for $j = 1, \dotsc, n$;
		\item $A \in \RR^{m \times n}$ with $A\T A = I_n$. 
	\end{enumerate}

\item Let $a, b \in \RR$ with $a < b$. Compute the Moore-Penrose inverse of the operator $A \colon L^2([a, b]) \to \RR$ with $Au = \int_a^b u(x) \dd{x}$. 
	\end{enumerate}
\end{question}

\begin{solution}
	\begin{enumerate}[(a)]
		\item \begin{enumerate}[(i)]
			\item Clearly $\Ran(A) = \RR$ and $\Nul(A)\P = \Span\qty{\vec e}$ where $\vec e$ is the all-ones vector. So $A\D$ must map $x \in \RR$ to $\vec e/n$, and therefore we have
			\[
			A\D = \vec e/n \in \RR^{n \times 1}.
			\]

			\item Clearly we have $\Ran(A) = \Span\qty{\vec e_j \mid a_j \neq 0} = \Nul(A)\P$ while $\Ran(A)\P = \Span\qty{\vec e_j \mid a_j = 0} = \Nul(A)$.

			It is easily seen that
			\[
			A\D = \diag(b_1, \dotsc, b_n) \in \RR^{n \times n}, \quad\text{where } b_i = \begin{cases}
				a_i^{-1} &\text{if $a_i \neq 0$}; \\
				0 &\text{if $a_i = 0$}. 
			\end{cases}
			\]

			\item If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal. In particular, $A$ is injective (so $\Nul(A) = \qty{0}$) and $\Ran(A) = \Span\qty{\vec u_1, \dotsc, \vec u_n}$. Extend $\vec u_1, \dotsc, \vec u_n$ to an orthonormal basis $\vec u_1, \dotsc, \vec u_m$ of $\RR^m$, then we must have
			\[
			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
			\]
			From this expression it is easily seen that $A\D = A\T$. 
		\end{enumerate}

	\item Clearly we have $\Ran(A) = \RR$ while $\Nul(A) = \qty{u \mid \int_a^b u(x) \dd{x} = 0}$. It is also easily seen that 
	\[
	\Nul(A)\P = \qty{v \mid \int_a^b v(x) u(x) \dd{x} = 0 \text{ if $\int_a^b u(x) \dd{x} = 0$}} = \Span\qty{1}. 
	\]

	Therefore we simply have that $A\D$ maps a constant $c \in \RR$ to the constant function $\frac{c}{b-a}$. 
	\end{enumerate}
\end{solution}

\begin{question}
	Many forward problems are either modelled as convolutions or they are modelled as the composition of several components, one of which is a convolution. Therefore convolutions play an important role in inverse problems. As in Exercise 1, let $\Omega =[0, 1]^2$ be the unit square and let $\XX = L^2(\Omega)$. A convolution is the special case of an integral operator $A \colon \XX \to \XX$ where the kernel has a simple structure: 
	\[
	(Au)(y) \ceq \int_\Omega k(y - x) u(x) \dd{x},
	\]
	for $k \in L^2(\Omega)$. It follows easily from Exercise 1 that $A$ is linear and bounded. 
	\begin{enumerate}[(a)]
		\item Although shown in general in Exercise 1, give an explicit form for the adjoint of the convolution. 
		\item Let $f = Au$. It follows from the convolution theorem that a convolution can be inverted by means of the Fourier transform
		\begin{equation} \label{eq:fourier}
		u = (2\pi)^{-\frac n2} \FF^{-1} \qty(\frac{\FF(f)}{\FF(k)}),
		\end{equation}
		where $\FF$ is the Fourier transform and $\FF^{-1}$ its inverse. Implement this formula in \emph{MATLAB} to deblur the blurry tree image $f$ generated by the script \verb|ex4b_generate_data.m|. Note that the script also outputs $\FF(k)$. Add some noise to the data and show that the inversion formula is ill-conditioned.

		Hint: Make use of the \emph{MATLAB} commands \verb|fft2| and \verb|ifft2|.

		\item Reformulate \cref{eq:fourier} so that the denominator is non-negative and give a stable approximation of this formula. Implement this formula in MATLAB and empirically show that it is stable.

		Hint: Make use of the \emph{MATLAB} command \verb|conj|. 
	\end{enumerate}
\end{question}

\begin{proof}
	\begin{enumerate}[(a)]
		\item We have by question 1 that
		\[
		(A^* v)(y) = \int_\Omega k(x - y)v(x) \dd{x}.
		\]
		\item 
	\end{enumerate}
\end{proof}
\end{document}

### 2020-10-27 modern_statistical_methods/msm_ES1.tex
worked on some example sheets, updated msm and topics summaries (todo: dist and inv prob)

@@

\begin{proof}
	Without loss of generality we can assume $\HH_1 = \HH$, where $\HH$ is the RKHS constructed in the proof of Theorem 5, so $\HH$ is the completion of $\Span\qty{k(\cdot, x) \mid x \in \XX}$. Now we can embed $\HH$ into $\HH_2$ by mapping $k(\cdot, x)$ to $k_x$. Clearly this mapping respects the inner product by construction, so this is a linear embedding. Therefore, we may assume $\HH_1 \subseteq \HH_2$.

	Now suppose $\HH_1 \subseteq \HH_2$ and decompose each $f \in \HH_2$ as $f = u + v$ with $u \in \HH_1$, $v \in \HH_1\P$. By the previous question we know that for every $x$ there exists a unique $k_x$ with $\ang{k_x, f} = f(x)$ for all $f$, and clearly we have $k_x \in \HH_1$.

	Then for each $f \in \HH_2, x \in \XX$ we have (since $k_x \in \HH_1$) that
	\[
	f(x) = \ang{k_x, f} = \ang{k_x, u} + \ang{k_x, v} = \ang{k_x, u} = u(x),
	\]
	which implies $v = 0$. Since this is the case for every $f \in \HH_2$, we conclude that $\HH_1\P = 0$ and therefore $\HH_1 = \HH_2$ (since Hilbert spaces are closed). 
\end{proof}
\end{document}

>>>>

¨

@@

\begin{proof}
%	Without loss of generality we can assume $\HH_1 = \HH$, where $\HH$ is the RKHS constructed in the proof of Theorem 5, so $\HH$ is the completion of $\Span\qty{k(\cdot, x) \mid x \in \XX}$. Now we can embed $\HH$ into $\HH_2$ by mapping $k(\cdot, x)$ to $k_x$. Clearly this mapping respects the inner product by construction, so this is a linear embedding. Therefore, we may assume $\HH_1 \subseteq \HH_2$. 
%	
%	Now suppose $\HH_1 \subseteq \HH_2$ and decompose each $f \in \HH_2$ as $f = u + v$ with $u \in \HH_1$, $v \in \HH_1\P$. By the previous question we know that for every $x$ there exists a unique $k_x$ with $\ang{k_x, f} = f(x)$ for all $f$, and clearly we have $k_x \in \HH_1$. 
%	
%	Then for each $f \in \HH_2, x \in \XX$ we have (since $k_x \in \HH_1$) that
%	\[
%	f(x) = \ang{k_x, f} = \ang{k_x, u} + \ang{k_x, v} = \ang{k_x, u} = u(x),
%	\]
%	which implies $v = 0$. Since this is the case for every $f \in \HH_2$, we conclude that $\HH_1\P = 0$ and therefore $\HH_1 = \HH_2$ (since Hilbert spaces are closed). 
Let $U = \HH_1 \cap \HH_2$. By the previous question, we know that $V$ must contain all functions of the form $k(\cdot, x)$. Now, take any $f \in \HH_1$ and write $f = u + v$ with $u \in U$, $v \in U\P$. Then we have
\[
v(x) = \ang{v, k(\cdot, x)} = 0 \quad\text{for all $x$}, 
\]
so $v = 0$. Therefore we find $f \in V$, so in particular we have $\HH_1 \subseteq V$. Analogously we have $\HH_2 \subseteq V$, which shows $\HH_1 = \HH_2$.
\end{proof}
\end{document}
